GSE158422: DNA methylation in Lung squamous cell carcinoma - gene methylation enrichment analysis - sensitivity analysis

### 2024-04-09 figs/GSE158422_sensitivity.Rmd
source
added analysis of number of probes

@@

```{r,al_down1,fig.height=10}

@@

@@

```{r,gsamethdownsample,fig.height=10}

>>>>

Do significant gene sets bias toward genes with more probes?

```{r,lam_genelength}

genes <- unique(gt$gene)
probecount <- unlist(lapply(genes,function(gene) { length(which(gt$gene==gene)) } ) )
names(probecount) <- genes
head(probecount)

sigsets <- rownames(lamsig)
nonsigsets <- rownames(subset(lamres,p.adjustANOVA>0.05))

sigmeancounts <- lapply(sigsets, function(set) {
  genes <- gs_symbols[[which(names(gs_symbols) ==set)]]
  counts <- probecount[which(names(probecount) %in% genes)]
  return(mean(counts))
} )

names(sigmeancounts) <- sigsets
sigmeancounts <- unlist(sigmeancounts)
summary(sigmeancounts)

nonsigmeancounts <- lapply(nonsigsets, function(set) {
  genes <- gs_symbols[[which(names(gs_symbols) ==set)]]
  counts <- probecount[which(names(probecount) %in% genes)]
  return(mean(counts))
} )

names(nonsigmeancounts) <- nonsigsets
nonsigmeancounts <- unlist(nonsigmeancounts)
summary(nonsigmeancounts)

mylist <- list("FDR>0.05"=nonsigmeancounts,"FDR<0.05"=sigmeancounts)

boxplot(mylist,col="white",ylab="probes per gene",cex=0,main="LAM",ylim=c(0,100)) ; grid()
beeswarm(mylist,add=TRUE,pch=1,cex=0.5)

par(mar = c(5.1, 4.1, 4.1, 2.1) )
pdf("fig3e.pdf",height=6,width=7)
boxplot(mylist,col="white",ylab="probes per gene",cex=0,main="LAM",ylim=c(0,100)) ; grid()
beeswarm(mylist,add=TRUE,pch=1,cex=0.5)
dev.off()

@@

```{r,al_down1,fig.height=10,eval=FALSE}

@@

Are gsameth significant pathways influenced by number of probes per gene?

```{r,gsa_genelength}

head(probecount)

allsets <- sub(" UP$","",rownames(gsaup))

gsaup_sets <- gsub(" UP$","",rownames(subset(gsaup,FDR<0.05)))
gsadn_sets <- gsub(" DN$","",rownames(subset(gsadn,FDR<0.05)))
sigsets <- unique(c(gsaup_sets,gsadn_sets))
length(sigsets)

nonsigsets <- setdiff(allsets,sigsets)

sigmeancounts <- lapply(sigsets, function(set) {
  genes <- gs_symbols[[which(names(gs_symbols) ==set)]]
  counts <- probecount[which(names(probecount) %in% genes)]
  return(mean(counts))
} )

names(sigmeancounts) <- sigsets
sigmeancounts <- unlist(sigmeancounts)
summary(sigmeancounts)

nonsigmeancounts <- lapply(nonsigsets, function(set) {
  genes <- gs_symbols[[which(names(gs_symbols) ==set)]]
  counts <- probecount[which(names(probecount) %in% genes)]
  return(mean(counts))
} )

names(nonsigmeancounts) <- nonsigsets
nonsigmeancounts <- unlist(nonsigmeancounts)
summary(nonsigmeancounts)

mylist <- list("FDR>0.05"=nonsigmeancounts,"FDR<0.05"=sigmeancounts)

boxplot(mylist,col="white",ylab="probes per gene",cex=0,main="GSA",ylim=c(0,100)) ; grid()
beeswarm(mylist,add=TRUE,pch=1,cex=0.5)

par(mar = c(5.1, 4.1, 4.1, 2.1) )
pdf("fig3f.pdf",height=6,width=7)
boxplot(mylist,col="white",ylab="probes per gene",cex=0,main="GSA",ylim=c(0,100)) ; grid()
beeswarm(mylist,add=TRUE,pch=1,cex=0.5)
dev.off()

@@

```{r,gsamethdownsample,fig.height=10,eval=FALSE}
